---
title: "Analyses and Graphs for Gibson et al. - Adaptation to novel hosts"
author: "Amanda Gibson"
date: "January 23, 2020"
output: html_document
---

Packages
```{r}
library(ggplot2)
library(MASS)
library(lme4)
library(ggthemes)
library(reshape)

#overdispersion function
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}  
```

Data sets
```{r}
# data set for sympatric hosts
d<-read.csv("data_Gibsonetal_focal_host_data.csv",header=TRUE)
d$Line<-factor(d$Line)
d$Rep<-factor(d$Rep)
summary(d)
# metadata for spreadsheet: A=date of assay, B=experimental evolution treatment of parasite; C=replicate parasite lineages within treatment; D=Host genotype assayed; E=technical replicate for assay; F= person counting, G=Number of live worms on food; H=Number of live worms on border between lawns; I=Total number of live hosts (column G+H); J=Total number of worms added, estimated from rows "worm counts"; K=Mortality rate, 1-(G+H)/I

# data for killing of N2 hosts
dN <-subset(d,d$Treatment!="worm counts"& d$Host=="N2")
summary(dN)
dN<- within(dN, Treatment <- relevel(Treatment, ref = "ancestor"))

# data for killing of CF3-wt-inv hosts = LTM1
dC <-subset(d,d$Treatment!="worm counts"& d$Host=="CF3-wt-inv")
summary(dC)
dC<- within(dC, Treatment <- relevel(Treatment, ref = "ancestor"))

# data for killing of novel hosts, JU1395 hosts
dJ <- read.csv("data_Gibsonetal_novel_host_data.csv",header=TRUE)
# metadata for spreadsheet: A=date of assay, B=experimental evolution treatment of parasite; C=replicate parasite lineage within treamtent; D=Host genotype assayed; E=technical replicate for assay; F= person counting, G=Number of live worms on food; H=Number of live worms on border between lawns; I=Total number of live hosts (column G+H);  J=Total number of worms added, estimated from rows "worm counts";K=Mortality rate, 1-(G+H)/I
dJ <-subset(dJ,dJ$Treatment!="worm counts")
head(dJ)
dJ$Line<-factor(dJ$Line)
dJ$Rep<-factor(dJ$Rep)
summary(dJ)

```


Adaptation to sympatric hosts: analysis. host line = N2; Table 1A, Table 2A, Table S1
```{r}
# Relevant numbers
# survival
n_number_live1<-tapply(dN$Live.total,list(dN$Treatment,dN$Line),mean)
n_number_live2<-melt(n_number_live1)
colnames(n_number_live2)=c("Treatment","Line","Live.total")
n_number_live<-tapply(n_number_live2$Live.total,list(n_number_live2$Treatment),mean,na.rm=T)
n_number_se<-tapply(n_number_live2$Live.total,list(n_number_live2$Treatment),sd,na.rm=T)/tapply(n_number_live2$Live.total,list(n_number_live2$Treatment),length)

#mortality
n_mortality1<-tapply(dN$Mortality,list(dN$Treatment,dN$Line),mean)
n_mortality2<-melt(n_mortality1)
colnames(n_mortality2)=c("Treatment","Line","Mortality")
n_mortality<-tapply(n_mortality2$Mortality,list(n_mortality2$Treatment),mean,na.rm=T)
n_mortality_se<-tapply(n_mortality2$Mortality,list(n_mortality2$Treatment),sd,na.rm=T)/tapply(n_mortality2$Mortality,list(n_mortality2$Treatment),length)

# SUMMARY OF SURVIVAL OF N2 HOSTS ACROSS PARASITE TREATMENTS - Table 1A
N2_survival<-cbind(n_number_live,n_number_se)
colnames(N2_survival)=c("Mean number surviving hosts","Standard error")
N2_survival

# SUMMARY OF MORTALITY OF N2 HOSTS ACROSS PARASITE TREATMENTS - Table 1A
N2_mortality<-cbind(n_mortality,n_mortality_se)
colnames(N2_mortality)=c("Mean mortality rate","Standard error")
N2_mortality

# conduct further analyses without ancestor
# exclude ancestor
dN <-subset(dN,dN$Treatment!="ancestor")
dN<- within(dN, Treatment <- relevel(Treatment, ref = "0-0"))

# fitting various models to survival counts for each replicate
# excluding offset(log(Total)) - total worms added same for all treatments; results identical with and without offset, except for intercept
m1<-glmer(dN$Live.total~Treatment+(1|Line),data=dN,family="poisson") # poisson
overdisp_fun(m1) # overdispersed

# negative binomial
m2<-glmer.nb(dN$Live.total~Treatment+(1|Line),data=dN) # negative binomial

# comparison - m2 better
stat<-2 * (logLik(m2) - logLik(m1))
pv<- pchisq(stat, df = 5 - 4, lower.tail = FALSE)
anova(m2,m1)

# TABLE 2A
# does experimental evolution treatment explain variation in survival?
m3 <- update(m2, . ~ . - Treatment)
anova(m3, m2) # yes

# results of m2, 0-0 is reference - TABLE S1
summary(m2)
(est <- cbind(Estimate = coef(m2), confint(m2)))

# results of m2, 100-0 as reference
dN<- within(dN, Treatment <- relevel(Treatment, ref = "100-0"))
m2_b<-glmer.nb(dN$Live.total~Treatment+(1|Line),data=dN) 
summary(m2_b)
(est <- cbind(Estimate = coef(m2_b), confint(m2_b)))

# results of m2, 50-50 as reference
dN<- within(dN, Treatment <- relevel(Treatment, ref = "50-50"))
m2_c<-glmer.nb(dN$Live.total~Treatment+(1|Line),data=dN)
summary(m2_c)
(est <- cbind(Estimate = coef(m2_c), confint(m2_c)))

# halve number of N2 hosts - comparison to JU1395 analysis
m2_halved<-glmer.nb(round(dN$Live.total/2,digits=0)~Treatment+(1|Line),data=dN)
m3_halved <- update(m2_halved, . ~ . - Treatment)
anova(m2_halved, m3_halved) # result the same

summary(m2_halved) # results the same

```

Adaptation to sympatric hosts: graph. host line = N2. Figure 2A

```{r}
dN$Treatment=factor(dN$Treatment,levels=c("0-0","0-100","100-0", "50-50"))

p2A <- ggplot(dN, aes(x=Treatment, y=Mortality)) +   
  geom_boxplot(outlier.colour="white",outlier.size=0,fill=c("white","#d9d9d9","#d9d9d9","#737373"),outlier.shape=8,lwd=1.2,color="gray10")+
  geom_jitter(shape=16,position=position_jitter(0.2),size=3,colour="gray10")+
  labs(x="")+
  labs(y="")+
  scale_x_discrete(labels=c("ancestor" = "Ancestor", "0-0" = "Control", "100-0"="100% N2","50-50"="50%","0-100"="100% CF3"))+
  scale_y_continuous(limits=c(0.4,1),breaks=c(0.4,0.5,0.6,0.7,0.8,0.9,1))+
  theme_base()+
  theme(axis.text.x = element_text(size=24))+
  theme(axis.text.y = element_text(size=24))+
  geom_vline(xintercept=1.5, linetype="dashed",color="gray30",size=1)+
  geom_vline(xintercept=3.5, linetype="dashed",color="gray30",size=1)

#FIGURE 2A
p2A

```

Adaptation to sympatric hosts: analysis. host line = CF3-wt-inv, Table 1B, Table 2B, table S2

```{r}
# Relevant numbers
# survival
c_number_live1<-tapply(dC$Live.total,list(dC$Treatment,dC$Line),mean)
c_number_live2<-melt(c_number_live1)
colnames(c_number_live2)=c("Treatment","Line","Live.total")
c_number_live<-tapply(c_number_live2$Live.total,list(c_number_live2$Treatment),mean,na.rm=T)
c_number_se<-tapply(c_number_live2$Live.total,list(c_number_live2$Treatment),sd,na.rm=T)/tapply(c_number_live2$Live.total,list(c_number_live2$Treatment),length)

#mortality
c_mortality1<-tapply(dC$Mortality,list(dC$Treatment,dC$Line),mean)
c_mortality2<-melt(c_mortality1)
colnames(c_mortality2)=c("Treatment","Line","Mortality")
c_mortality<-tapply(c_mortality2$Mortality,list(c_mortality2$Treatment),mean,na.rm=T)
c_mortality_se<-tapply(c_mortality2$Mortality,list(c_mortality2$Treatment),sd,na.rm=T)/tapply(c_mortality2$Mortality,list(c_mortality2$Treatment),length)

# SUMMARY OF SURVIVAL OF CF3-wt-inv HOSTS ACROSS PARASITE TREATMENTS - Table 1B
CF3_survival<-cbind(c_number_live,c_number_se)
colnames(CF3_survival)=c("Mean number surviving hosts","Standard error")
CF3_survival

# SUMMARY OF MORTALITY OF CF3-wt-inv  HOSTS ACROSS PARASITE TREATMENTS- Table 1B
CF3_mortality<-cbind(c_mortality,c_mortality_se)
colnames(CF3_mortality)=c("Mean mortality rate","Standard error")
CF3_mortality

# conduct further analyses without ancestor
# exclude ancestor
dC <-subset(dC,dC$Treatment!="ancestor")
dC <- within(dC, Treatment <- relevel(Treatment, ref = "0-0"))

# fitting various models to survival counts for each replicate
# excluding offset(log(Total)) - total worms added same for all treatments; results identical with and without offset, except for intercept
m1<-glmer(dC$Live.total~Treatment+(1|Line),data=dC,family="poisson") # poisson
overdisp_fun(m1) # overdispersed

m2<-glmer.nb(dC$Live.total~Treatment+(1|Line),data=dC) 
m3<-glm.nb(dC$Live.total~Treatment+Line,data=dC) 
# negative binomial - fitting with Line as random or fixed effect has no effect on results, but you do need line
anova(m3,test="Chisq")
m4<-glm.nb(dC$Live.total~Treatment,data=dC) #line out entirely

# comparison - m2 better
stat<-2 * (logLik(m2) - logLik(m1))
pv<- pchisq(stat, df = 5 - 4, lower.tail = FALSE)
anova(m2,m1)

# TABLE 2B
# does experimental evolution treatment explain variation in survival?
m3 <- update(m2, . ~ . - Treatment)
anova(m2, m3) # no
summary(m2) # Table S2

# halve number of N2 hosts - comparison to JU1395 analysis
m2_halved<-glmer.nb(round(dC$Live.total/2,digits=0)~Treatment+(1|Line),data=dC)
m3_halved <- update(m2_halved, . ~ . - Treatment)
anova(m2_halved, m3_halved) # result the same

summary(m2_halved) # results the same
```

Adaptation to sympatric hosts: graph. host line = CF3-wt-inv. Figure 2B
```{r}
dC$Treatment=factor(dC$Treatment,levels=c("0-0","0-100","100-0","50-50"))

p2B <- ggplot(dC, aes(x=Treatment, y=Mortality)) +   
  geom_boxplot(outlier.colour="white",outlier.size=0,fill=c("white","#d9d9d9","#d9d9d9","#737373"),outlier.shape=8,lwd=1.2,color="gray10")+
  geom_jitter(shape=16,position=position_jitter(0.2),size=3,colour="gray10")+
  labs(x="")+
  labs(y="")+
  scale_x_discrete(labels=c("ancestor" = "Ancestor", "0-0" = "Control", "100-0"="100% N2","50-50"="50%","0-100"="100% CF3"))+
  scale_y_continuous(limits=c(0.4,1),breaks=c(0.4,0.5,0.6,0.7,0.8,0.9,1))+
  theme_base()+
  theme(axis.text.x = element_text(size=24))+
  theme(axis.text.y = element_text(size=24))+
    geom_vline(xintercept=1.5, linetype="dashed",color="gray30",size=1)+
  geom_vline(xintercept=3.5, linetype="dashed",color="gray30",size=1)


#FIGURE 2B
p2B
```

Adaptation to novel hosts: analysis. host line = JU1395; Table 1C, Table 2C, Table

```{r}
# Relevant numbers
# survival
j_number_live1<-tapply(dJ$Live.total,list(dJ$Treatment,dJ$Line),mean)
j_number_live2<-melt(j_number_live1)
colnames(j_number_live2)=c("Treatment","Line","Live.total")
j_number_live<-tapply(j_number_live2$Live.total,list(j_number_live2$Treatment),mean,na.rm=T)
j_number_se<-tapply(j_number_live2$Live.total,list(j_number_live2$Treatment),sd,na.rm=T)/tapply(j_number_live2$Live.total,list(j_number_live2$Treatment),length)

#mortality
j_mortality1<-tapply(dJ$Mortality,list(dJ$Treatment,dJ$Line),mean)
j_mortality2<-melt(j_mortality1)
colnames(j_mortality2)=c("Treatment","Line","Mortality")
j_mortality<-tapply(j_mortality2$Mortality,list(j_mortality2$Treatment),mean,na.rm=T)
j_mortality_se<-tapply(j_mortality2$Mortality,list(j_mortality2$Treatment),sd,na.rm=T)/tapply(j_mortality2$Mortality,list(j_mortality2$Treatment),length)

# SUMMARY OF SURVIVAL OF JU1395 HOSTS ACROSS PARASITE TREATMENTS
JU_survival<-cbind(j_number_live,j_number_se)
colnames(JU_survival)=c("Mean number surviving hosts","Standard error")
JU_survival

# SUMMARY OF MORTALITY OF JU1395  HOSTS ACROSS PARASITE TREATMENTS
JU_mortality<-cbind(j_mortality,j_mortality_se)
colnames(JU_mortality)=c("Mean mortality rate","Standard error")
JU_mortality

# conduct further analyses without ancestor
# exclude ancestor
dJ <-subset(dJ,dJ$Treatment!="Ancestor")
dJ<- within(dJ, Treatment <- relevel(Treatment, ref = "0-0"))

# fitting various models to survival counts for each replicate
# excluding offset(log(Total)) - total worms added same for all treatments; results identical with and without offset, except for intercept
m1<-glmer(dJ$Live.total~Treatment+(1|Line),data=dJ,family="poisson") # poisson
overdisp_fun(m1) # overdispersed
m2<-glmer.nb(dJ$Live.total~Treatment+(1|Line),data=dJ) # negative binomial

# comparison - m2 better
stat<-2 * (logLik(m2) - logLik(m1))
pv<- pchisq(stat, df = 5 - 4, lower.tail = FALSE)
anova(m2,m1)

#TABLE 1C
# does experimental evolution treatment explain variation in survival?
m3 <- update(m2, . ~ . - Treatment)
anova(m2, m3) # yes

# results of m2, 0-0 is reference
summary(m2)
(est <- cbind(Estimate = coef(m2), confint(m2)))

# results of m2, 100-0 as reference
dJ<- within(dJ, Treatment <- relevel(Treatment, ref = "100-0"))
m2_b<-glmer.nb(dJ$Live.total~Treatment+(1|Line),data=dJ) 
summary(m2_b)
(est <- cbind(Estimate = coef(m2_b), confint(m2_b)))

# results of m2, 50-50 as reference
dJ<- within(dJ, Treatment <- relevel(Treatment, ref = "50-50"))
m2_c<-glmer.nb(dJ$Live.total~Treatment+(1|Line),data=dJ)
summary(m2_c)
(est <- cbind(Estimate = coef(m2_c), confint(m2_c)))

# results of m2, 0-100 as reference
dJ<- within(dJ, Treatment <- relevel(Treatment, ref = "0-100"))
m2_c<-glmer.nb(dJ$Live.total~Treatment+(1|Line),data=dJ)
summary(m2_c)
(est <- cbind(Estimate = coef(m2_c), confint(m2_c)))

########### check
# reducing replicate number - this analysis has 8, relative to 4 replicates for N2 and LTM1
dJs<-subset(dJ,dJ$Rep!="5"&dJ$Rep!="6"&dJ$Rep!="7"&dJ$Rep!="8")
dJs$Rep=factor(dJs$Rep)

# fitting various models to survival counts for each replicate
# excluding offset(log(Total)) - total worms added same for all treatments; results identical with and without offset, except for intercept
m4<-glmer(dJs$Live.total~Treatment+(1|Line),data=dJs,family="poisson") # poisson
overdisp_fun(m4) # overdispersed
m5<-glmer.nb(dJs$Live.total~Treatment+(1|Line),data=dJs) # negative binomial

# comparison - m2 better
stat<-2 * (logLik(m5) - logLik(m4))
pv<- pchisq(stat, df = 5 - 4, lower.tail = FALSE)
anova(m5,m4)

# does experimental evolution treatment explain variation in survival?
m6 <- update(m5, . ~ . - Treatment)
anova(m5, m6) # yes

# results of m2, 0-0 is reference
summary(m5)
(est <- cbind(Estimate = coef(m5), confint(m5)))

# results of m5, 100-0 as reference
dJs<- within(dJs, Treatment <- relevel(Treatment, ref = "100-0"))
m5_b<-glmer.nb(dJs$Live.total~Treatment+(1|Line),data=dJs) 
summary(m5_b)
(est <- cbind(Estimate = coef(m5_b), confint(m5_b)))

# results of m5, 50-50 as reference
dJs<- within(dJs, Treatment <- relevel(Treatment, ref = "50-50"))
m5_c<-glmer.nb(dJs$Live.total~Treatment+(1|Line),data=dJs)
summary(m5_c)
(est <- cbind(Estimate = coef(m5_c), confint(m5_c)))

# results of m5, 0-100 as reference
dJs<- within(dJs, Treatment <- relevel(Treatment, ref = "0-100"))
m5_d<-glmer.nb(dJs$Live.total~Treatment+(1|Line),data=dJs)
summary(m5_d)
(est <- cbind(Estimate = coef(m5_d), confint(m5_d)))
```

Adaptation to novel host: graph. host line = JU1395. Figure 3

```{r}
dJ$Treatment=factor(dJ$Treatment,levels=c("0-0","0-100","100-0","50-50"))

p3 <- ggplot(dJ, aes(x=Treatment, y=Mortality)) +   
  geom_boxplot(outlier.colour="white",outlier.size=0,fill=c("white","#d9d9d9","#d9d9d9","#737373"),outlier.shape=8,lwd=1.2,color="gray10")+
  geom_jitter(shape=16,position=position_jitter(0.2),size=3,colour="gray10")+
  labs(x="")+
  labs(y="")+
  scale_x_discrete(labels=c("0-0" = "Control", "100-0"="100% N2","50-50"="50%","0-100"="100% CF3"))+
  scale_y_continuous(limits=c(0.4,1),breaks=c(0.4,0.5,0.6,0.7,0.8,0.9,1))+
  theme_base()+
  theme(axis.text.x = element_text(size=24))+
  theme(axis.text.y = element_text(size=24))+
  geom_vline(xintercept=1.5, linetype="dashed",color="gray30",size=1)+
  geom_vline(xintercept=3.5, linetype="dashed",color="gray30",size=1)

#FIGURE 3
p3

```

Variation across infection of novel hosts, Table S4
```{r}
# Total worms alive
dJmL=tapply(dJ$Live.total,list(dJ$Treatment,dJ$Line),mean)
dJL=melt(dJmL)
colnames(dJL)=c("Treatment","Line","Live.total")

dJm=tapply(dJL$Live.total,list(dJL$Treatment),mean)
dJs=tapply(dJL$Live.total,list(dJL$Treatment),sd)
dJcv=dJs/dJm

#95% CI
# bootstrap CV
m = 10000 

# 0-0
cvBoot00=rep(0,m)
dJ00=subset(dJ,dJ$Treatment=="0-0")
for (i in 1:m) { 
  dboot1<-sample(dJ00$Live.total[dJ00$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ00$Live.total[dJ00$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ00$Live.total[dJ00$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ00$Live.total[dJ00$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ00$Live.total[dJ00$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ00$Live.total[dJ00$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot00[i]<-dJs/dJm
}
cvBoot00 = sort(cvBoot00,decreasing=FALSE) 
low00<-cvBoot00[m*0.025]
high00<-cvBoot00[m*0.975] 

# 0-100
cvBoot0100=rep(0,m)
dJ0100=subset(dJ,dJ$Treatment=="0-100")
for (i in 1:m) { 
  dboot1<-sample(dJ0100$Live.total[dJ0100$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ0100$Live.total[dJ0100$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ0100$Live.total[dJ0100$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ0100$Live.total[dJ0100$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ0100$Live.total[dJ0100$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ0100$Live.total[dJ0100$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot0100[i]<-dJs/dJm
}
cvBoot0100 = sort(cvBoot0100,decreasing=FALSE) 
low0100<-cvBoot0100[m*0.025]
high0100<-cvBoot0100[m*0.975] 

# 50-50
cvBoot5050=rep(0,m)
dJ5050=subset(dJ,dJ$Treatment=="50-50")
for (i in 1:m) { 
  dboot1<-sample(dJ5050$Live.total[dJ5050$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ5050$Live.total[dJ5050$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ5050$Live.total[dJ5050$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ5050$Live.total[dJ5050$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ5050$Live.total[dJ5050$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ5050$Live.total[dJ5050$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot5050[i]<-dJs/dJm
}
cvBoot5050 = sort(cvBoot5050,decreasing=FALSE) 
low5050<-cvBoot5050[m*0.025]
high5050<-cvBoot5050[m*0.975] 

# 100-0
cvBoot1000=rep(0,m)
dJ1000=subset(dJ,dJ$Treatment=="100-0")
for (i in 1:m) { 
  dboot1<-sample(dJ1000$Live.total[dJ1000$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ1000$Live.total[dJ1000$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ1000$Live.total[dJ1000$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ1000$Live.total[dJ1000$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ1000$Live.total[dJ1000$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ1000$Live.total[dJ1000$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot1000[i]<-dJs/dJm
}
cvBoot1000 = sort(cvBoot1000,decreasing=FALSE) 
low1000<-cvBoot1000[m*0.025]
high1000<-cvBoot1000[m*0.975] 

# SUMMARY OF VARIANCE IN SURVIVAL COUNTS OF NOVEL HOSTS BY PARASITE TREATMENT
tCV_survival<-matrix(c(low00,dJcv[1],high00,low0100,dJcv[2],high0100,low1000,dJcv[3],high1000,low5050,dJcv[4],high5050),
            nrow=3,ncol=4)
colnames(tCV_survival)=c("0-0","0-100","100-0","50-50")
rownames(tCV_survival)=c("lower bound 95% CI","Coefficient of variation of Live.total rate","upper bound 95% CI")
tCV_survival

# Mortality
dJmL=tapply(dJ$Mortality,list(dJ$Treatment,dJ$Line),mean)
dJsL=tapply(dJ$Mortality,list(dJ$Treatment,dJ$Line),sd)/sqrt(8)
dJL=melt(dJmL)
colnames(dJL)=c("Treatment","Line","Mortality")

dJm=tapply(dJL$Mortality,list(dJL$Treatment),mean)
dJs=tapply(dJL$Mortality,list(dJL$Treatment),sd)
dJcv=dJs/dJm

# bootstrap CV
m = 10000 

# 0-0
cvBoot00=rep(0,m)
dJ00=subset(dJ,dJ$Treatment=="0-0")
for (i in 1:m) { 
  dboot1<-sample(dJ00$Mortality[dJ00$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ00$Mortality[dJ00$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ00$Mortality[dJ00$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ00$Mortality[dJ00$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ00$Mortality[dJ00$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ00$Mortality[dJ00$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot00[i]<-dJs/dJm
}
cvBoot00 = sort(cvBoot00,decreasing=FALSE) 
low00<-cvBoot00[m*0.025]
high00<-cvBoot00[m*0.975] 

# 0-100
cvBoot0100=rep(0,m)
dJ0100=subset(dJ,dJ$Treatment=="0-100")
for (i in 1:m) { 
  dboot1<-sample(dJ0100$Mortality[dJ0100$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ0100$Mortality[dJ0100$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ0100$Mortality[dJ0100$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ0100$Mortality[dJ0100$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ0100$Mortality[dJ0100$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ0100$Mortality[dJ0100$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot0100[i]<-dJs/dJm
}
cvBoot0100 = sort(cvBoot0100,decreasing=FALSE) 
low0100<-cvBoot0100[m*0.025]
high0100<-cvBoot0100[m*0.975] 

# 50-50
cvBoot5050=rep(0,m)
dJ5050=subset(dJ,dJ$Treatment=="50-50")
for (i in 1:m) { 
  dboot1<-sample(dJ5050$Mortality[dJ5050$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ5050$Mortality[dJ5050$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ5050$Mortality[dJ5050$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ5050$Mortality[dJ5050$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ5050$Mortality[dJ5050$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ5050$Mortality[dJ5050$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot5050[i]<-dJs/dJm
}
cvBoot5050 = sort(cvBoot5050,decreasing=FALSE) 
low5050<-cvBoot5050[m*0.025]
high5050<-cvBoot5050[m*0.975] 

# 100-0
cvBoot1000=rep(0,m)
dJ1000=subset(dJ,dJ$Treatment=="100-0")
for (i in 1:m) { 
  dboot1<-sample(dJ1000$Mortality[dJ1000$Line==1],8,replace=TRUE)
  dboot2<-sample(dJ1000$Mortality[dJ1000$Line==2],8,replace=TRUE)
  dboot3<-sample(dJ1000$Mortality[dJ1000$Line==3],8,replace=TRUE)
  dboot4<-sample(dJ1000$Mortality[dJ1000$Line==4],8,replace=TRUE)
  dboot5<-sample(dJ1000$Mortality[dJ1000$Line==5],8,replace=TRUE)
  dboot6<-sample(dJ1000$Mortality[dJ1000$Line==6],8,replace=TRUE)
  dJmL=sapply(list(dboot1,dboot2,dboot3,dboot4,dboot5,dboot6),mean)
  dJm=mean(dJmL)
  dJs=sd(dJmL)
  cvBoot1000[i]<-dJs/dJm
}
cvBoot1000 = sort(cvBoot1000,decreasing=FALSE) 
low1000<-cvBoot1000[m*0.025]
high1000<-cvBoot1000[m*0.975] 

# SUMMARY OF VARIANCE IN MORTALITY RATE OF NOVEL HOSTS BY PARASITE TREATMENT
tCV_mortality<-matrix(c(low00,dJcv[1],high00,low0100,dJcv[2],high0100,low1000,dJcv[3],high1000,low5050,dJcv[4],high5050),
            nrow=3,ncol=4)
colnames(tCV_mortality)=c("0-0","0-100","100-0","50-50")
rownames(tCV_mortality)=c("lower bound 95% CI","Coefficient of variation of mortality rate","upper bound 95% CI")
tCV_mortality

```
Plotting mortality estimates of ancestral Sm2170 on N2 and LTM1
```{r}
dS<-subset(d,d$Treatment=="ancestor")
dS$Host=factor(dS$Host,levels=c("N2","CF3-wt-inv"))
dSmeans=c(0,tapply(dS$Mortality,list(dS$Host),mean),0)
dSse=c(0,tapply(dS$Mortality,list(dS$Host),sd)/sqrt(4),0)
x=c(1,2,3,4)

par(lwd=2)
par(tck=0.02)
par(las=1)
par(mar=c(4,4,4,4))
plot(x=x,y=dSmeans,type="p",
     ylim=c(0.5,1),
     xaxt="n",
     xlab="",
     ylab="",
     cex.axis=2,
     pch=21,bg="tomato",col="gray20",cex=1,lwd=3)
segments(x,dSmeans-dSse,x,dSmeans+dSse,
       lwd=3,col="gray20")
points(x,dSmeans,pch=21,bg="firebrick2",col="gray20",cex=4,lwd=2)


```
